---
title: "Univariate"
output: github_document

---

## Overview


The univariate package creates ggplot graphs specialized in comparing model predictions with actual outcomes over a given variable.  This can be helpful in illustrating what predictive behavior a model is catching (or failing to catch), especially with regard to potentially-correlated variables. It can also be useful in determining if a dependent variable requires transformation to better capture the targeted behavior.

This plot function was designed primarily for use with logistic regression where the target is either a 1 or 0.

## Installation

The easiest way is to install from GitHub directly using devtools:

```{r download, eval = FALSE}
# install.packages("devtools")
devtools::install_github("SignaLtoNoise/univariate")
```


## Basic Usage

```{r setup, include = FALSE}
setwd("C:/Users/mwilcox/Desktop")
df <- read.csv('adult.csv')

df$HighIncomeFlag <- ifelse(df$highincome == ' >50K', 1, 0)

modelinit <- glm(HighIncomeFlag ~ education.num
                 + relationship
                 , data = df, family = 'binomial')
df$predict <- predict(modelinit, newdata = df, type = 'response')

```

`univariate()` uses lazyeval to interpret variable columns in the data frame. The first four arguments are required:

* `df`: the dataframe containing the variables
* `groupvar`: the X-axis variable name to plot actual vs prediction across
* `actual`: the targeted observed value column name (1 or 0)
* `model`: the predicted model result column name


Because of lazyeval, the three column variables must be wrapped in quotes.

```{r initial example, message=FALSE}
library(univariate)
univariate(df, "age", "HighIncomeFlag", "predict")
```

The modeled predictions are graphed as a red line, while the actual target is graphed as a black scatterplot. Underneath is plotted a histogram of the groupvar to better understand the underlying behavior.

`univariate()` will detect if the groupvar is a continous or a factor, and will change the plot accordingly:

```{r example factor}
# groupvar is a factor
univariate(df, "sex", "HighIncomeFlag", "predict")
```

## Secondary Features

### Rounding
On continuous variables, it is sometimes necessary to create rounded values to make a more comprehensible plot. The `roundvalue` argument sets the accuracy to round to.  The `roundfunc` argument defaults to "round", but can be additionally set to "floor" or "ceiling".

```{r example round, message=FALSE}
univariate(df, "fnlwgt", "HighIncomeFlag", "predict", roundvalue = 10000, roundfunc = 'floor')
```

### Range Limit
Similarly, it is sometimes necessary to truncate long tails so that the relevant range is given clarity. The `grouprange` argument defines the numeric range to plot along the `groupvar` variable:

```{r example range, message=FALSE}
univariate(df, "fnlwgt", "HighIncomeFlag", "predict", grouprange = c(0, 600000),
           roundvalue = 10000, roundfunc = 'floor')
```

By default, the entire range of `groupvar` will be plotted.  For factor variables, `grouprange` selects the values along the X-Axis corresponding to the range.

### Custom Title and Label Angles
Univariate will create a default title, but you can replace it by feeding a string to the `titlestring` argument.

If there are many factor variables, you can specify a degree to angle the labels from the X-axis using the `labelangle` argument.
```{r example title angle, message=FALSE}
univariate(df, "occupation", "HighIncomeFlag", "predict",
           labelangle = 20, titlestring = "Univariate Plot with 20-Degree Labels")
```


